---
title: "3CT"
output:
  pdf_document: default
  html_document: default
date: '2022-12-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}

# Loading
library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans)

df <- read_excel("3CT_larger_arena new_3weeks.xlsx") #this fiel for larger arena

```

Add 1cm zone with surrounding zone
```{r}
df <- df %>% mutate(social_time=(`In zone SocialZone multiCond cumulative s`+df$SocialZoneRound))

df <- df %>% mutate(novel_time=(`In zone NoveltyZone multiCond cumulative s`+df$NoveltyZoneRound))

```


Removal of mouse 224 from dataset it turned it it was not double transgenic
```{r}
# #df$ID<- as.numeric(df$ID) #Change mouse number from class character to class numeric
# df<-df[!(df$ID==429),]
# df<-df[!(df$ID==374),]
# df<-df[!(df$ID==461),]
# df<-df[!(df$ID==463),]
# df<-df[!(df$ID==402),]

#remove TG animals to check single TG behaviour
# 
df<-df[!(df$ID==423),]
df<-df[!(df$ID==376),]
df<-df[!(df$ID==372),]
df<-df[!(df$ID==398),]
df<-df[!(df$ID==400),]
df<-df[!(df$ID==459),]

```


We will primarily analyse time spent in the different zones. We expect:

Habituation phase: time spent in neutral, social, and novel zone will be equal.
For social phase (1 mouse introduced): time spent in social will be bigger than neutral and novel zone combined.
for novelty phase (1 novel mouse introduced into novelty zone): time spent in novelty zone will be bigger then in neutral and social zone combined.


Divide dataset up into Habituation, social, and novelty phases

```{r}
#habituation_cond <-  df_complete %>% filter(Condition == "Hab")
#social_cond<-  df_complete %>% filter(Condition == "Social")
#novelty_cond<-  df_complete %>% filter(Condition == "Novel")

#use this for LMM approach with full sample size
habituation_cond <-  df %>% filter(Condition == "Hab")
social_cond<-  df %>% filter(Condition == "Social")
novelty_cond<-  df %>% filter(Condition == "Novel")
```

with 1cm zone and surrounding zone
```{r}
habituation_cond <- habituation_cond %>% mutate(social_preference_relative=((`In zone SocialZone multiCond cumulative s`-`In zone NoveltyZone multiCond cumulative s`)/(`In zone NoveltyZone multiCond cumulative s`+`In zone SocialZone multiCond cumulative s`)))



social_cond <- social_cond %>% mutate(social_preference_relative=((social_cond$social_time-social_cond$novel_time)/(social_cond$social_time+social_cond$novel_time)))

novelty_cond <- novelty_cond %>% mutate(novelty_preference_relative=((novelty_cond$novel_time-novelty_cond$social_time)/(novelty_cond$novel_time+novelty_cond$social_time)))

```



LMM sociability

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

social_cond$Week <- factor(social_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(social_preference_relative ~  Week + (1|ID), data=social_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova,pairwise ~ Week, conf.level = 0.95)
emmeans_results

class(emmeans_results)

emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
class(emm_single)

pairwise_results <- contrast(emm_single, method = "pairwise")
class(pairwise_results)

#pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")

# Perform pairwise comparisons
#pairwise_results <- contrast(emmeans_results, method = "pairwise")
# Display the results
summary(pairwise_results)

#bonferroni
#pwc <- pairs(emmeans_results, adjust = "bon")
#pwc

pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
summary(pwc)
```





plot

```{r}
cbbPalette <- c("#fcfdbf", "#feca8d", "#D67236", "#f1605d", "#9e2f7f", "#440f76")

p_load(ggbeeswarm, ggsignif)

ggplot(social_cond, aes(x = Week, y = social_preference_relative)) +
  geom_beeswarm(data = social_cond, size = 2.5, aes(color = ID)) +
  ylab("Social Preference") +
  xlab("Timepoint") +
  # geom_signif(comparisons = list(c("Preinduction", "3rd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 1.3,
  #             annotations = c("1.0000")) +
  # geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 1.15,
  #             annotations = c(" 0.0477")) +
  # geom_signif(comparisons = list(c("3rd week post induction", "2nd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 1,
  #             annotations = c("0.1927")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = social_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))

```
```{r}

levels(social_cond$Week)[levels(social_cond$Week) == "1st week post induction"] <- "1st week"
levels(social_cond$Week)[levels(social_cond$Week) == "2nd week post induction"] <- "2nd week"

ggplot(social_cond, aes(x = Week, y = social_preference_relative)) +
  geom_beeswarm(data = social_cond, size = 2.5, aes(color = ID)) +
  ylab("Social Preference") +
  xlab("Timepoint") +
  # geom_signif(comparisons = list(c("Preinduction", "1st week")),
  #             map_signif_level = TRUE,
  #             y_position = 1.3,
  #             annotations = c("ns")) +
  # geom_signif(comparisons = list(c("Preinduction", "2nd week")),
  #             map_signif_level = TRUE,
  #             y_position = 1.15,
  #             annotations = c("**")) +
  # geom_signif(comparisons = list(c("1st week", "2nd week")),
  #             map_signif_level = TRUE,
  #             y_position = 1,
  #             annotations = c("ns")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = social_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 16))
```



social novelty

LLM

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

novelty_cond$Week <- factor(novelty_cond$Week , levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(novelty_preference_relative ~  Week + (1|ID), data=novelty_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc
```


plot

```{r}
cbbPalette <- c("#fcfdbf", "#feca8d", "#D67236", "#f1605d", "#9e2f7f", "#440f76")

p_load(ggbeeswarm, ggsignif)

ggplot(novelty_cond, aes(x = Week, y = novelty_preference_relative)) +
  geom_beeswarm(data = novelty_cond, size = 2.5, aes(color = ID)) +
  ylab("Novelty Preference") +
  xlab("Timepoint") +
  # geom_signif(comparisons = list(c("Preinduction", "3rd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 1.3,
  #             annotations = c("1.0000")) +
  # geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 1.15,
  #             annotations = c(" 1.0000")) +
  # geom_signif(comparisons = list(c("3rd week post induction", "2nd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 1,
  #             annotations = c("1.0000")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = novelty_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))

```

```{r}

levels(novelty_cond$Week)[levels(novelty_cond$Week) == "1st week post induction"] <- "1st week"
levels(novelty_cond$Week)[levels(novelty_cond$Week) == "2nd week post induction"] <- "2nd week"

ggplot(novelty_cond, aes(x = Week, y = novelty_preference_relative)) +
  geom_beeswarm(data = novelty_cond, size = 2.5, aes(color = ID)) +
  ylab("Novelty Preference") +
  xlab("Timepoint") +
  # geom_signif(comparisons = list(c("Preinduction", "1st week")),
  #             map_signif_level = TRUE,
  #             y_position = 1.3,
  #             annotations = c("ns")) +
  # geom_signif(comparisons = list(c("Preinduction", "2nd week")),
  #             map_signif_level = TRUE,
  #             y_position = 1.15,
  #             annotations = c("ns")) +
  # geom_signif(comparisons = list(c("1st week", "2nd week")),
  #             map_signif_level = TRUE,
  #             y_position = 1,
  #             annotations = c("ns")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = novelty_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 16))
```


# T-test

Data set where participants not completing all data collection time points are sorted out
```{r}
# Filter out so data is only for the first two weeks
TwoWeeks_df <- df %>% filter(Week == "Preinduction" | Week == "1st week post induction")

```

#Filter out subjects that did not complete
```{r}

# Determine the weeks each participant should have
required_weeks <- TwoWeeks_df %>% 
  distinct(Week) %>% 
  pull()

# Count the number of weeks each participant has data for
participant_week_counts <- TwoWeeks_df %>%
  group_by(ID) %>%
  summarize(week_count = n_distinct(Week))

# Identify participants with data for all required weeks
complete_participants <- participant_week_counts %>%
  filter(week_count == length(required_weeks)) %>%
  pull(ID)

# Filter the dataset to include only complete participants
df_complete <- TwoWeeks_df %>%
  filter(ID %in% complete_participants)

# View the new dataset
head(df_complete)
```

```{r}
df_complete$ID <- as.factor(df_complete$ID)
df_complete$Week <- factor(df_complete$Week , levels = c("Preinduction","1st week post induction"))

```



Divide dataset up into Habituation, social, and novelty phases

```{r}
habituation_cond <-  df_complete %>% filter(Condition == "Hab")
social_cond<-  df_complete %>% filter(Condition == "Social")
novelty_cond<-  df_complete %>% filter(Condition == "Novel")

```


```{r}

habituation_cond <- habituation_cond %>% mutate(social_preference_relative=((`In zone SocialZone cumulative s`-`In zone NoveltyZone cumulative s`)/(`In zone NoveltyZone cumulative s`+`In zone SocialZone cumulative s`)))


#As described in Baronio 2015 
#https://journals.plos.org/plosone/article/file?id=10.1371/journal.pone.0116363&type=printable 

social_cond <- social_cond %>% mutate(social_preference_relative=((`In zone SocialZone cumulative s`-`In zone NoveltyZone cumulative s`)/(`In zone NoveltyZone cumulative s`+`In zone SocialZone cumulative s`)))

novelty_cond <- novelty_cond %>% mutate(novelty_preference_relative=((`In zone NoveltyZone cumulative s`-`In zone SocialZone cumulative s`)/(`In zone SocialZone cumulative s`+`In zone NoveltyZone cumulative s`)))

```

## Social preference 
```{r}

by(social_cond$social_preference_relative, social_cond$Week, shapiro.test)

# Perform paired t-test
t_test_result <- t.test(social_preference_relative ~ Week, data = social_cond, paired = TRUE)

# Print the result of the t-test
print(t_test_result)

# Extract the p-value from the test result
p_value <- t_test_result$p.value
```

```{r}
# Function to convert p-value to significance annotations
get_significance_label <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}
```


```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
significance_label <- get_significance_label(p_value)

ggplot(social_cond, aes(x = Week, y = social_preference_relative)) +
  geom_beeswarm(data = social_cond, size = 2.5, aes(color = ID)) +
  ylab("social preference") +
  xlab("Timepoint") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              annotations = significance_label) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = social_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```


## novelty preference 
```{r}

by(novelty_cond$novelty_preference_relative, novelty_cond$Week, shapiro.test)

# Perform paired t-test
t_test_result <- t.test(novelty_preference_relative ~ Week, data = novelty_cond, paired = TRUE)

# Print the result of the t-test
print(t_test_result)

# Extract the p-value from the test result
p_value <- t_test_result$p.value
```

```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
significance_label <- get_significance_label(p_value)

ggplot(novelty_cond, aes(x = Week, y = novelty_preference_relative)) +
  geom_beeswarm(data = novelty_cond, size = 2.5, aes(color = ID)) +
  ylab("Novelty preference") +
  xlab("Timepoint") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              annotations = significance_label) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = novelty_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```


#habituation

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

habituation_cond$Week <- factor(habituation_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction"))
habituation_cond$distance = habituation_cond$`Distance moved cm`
#creating a linear model of the same variables as included in the anova
m_anova <-lmer(habituation_cond$distance ~  Week + (1|ID), data=habituation_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova,pairwise ~ Week, conf.level = 0.95)
emmeans_results

class(emmeans_results)

emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
class(emm_single)

pairwise_results <- contrast(emm_single, method = "pairwise")
class(pairwise_results)

#pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")

# Perform pairwise comparisons
#pairwise_results <- contrast(emmeans_results, method = "pairwise")
# Display the results
summary(pairwise_results)

#bonferroni
#pwc <- pairs(emmeans_results, adjust = "bon")
#pwc

pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
summary(pwc)
```


```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
#significance_label <- get_significance_label(p_value)

ggplot(habituation_cond, aes(x = Week, y = distance)) +
  geom_beeswarm(data = habituation_cond, size = 2.5, aes(color = ID)) +
  ylab("Distance (cm)") +
  xlab("Timepoint habituation condition") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              y_position = 4600,
              annotations = c("ns")) +
  geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 4350,
              annotations = c("*")) +
  geom_signif(comparisons = list(c("1st week post induction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 4100,
              annotations = c("*")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = habituation_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

social_cond$Week <- factor(social_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction"))
social_cond$distance = social_cond$`Distance moved cm`
#creating a linear model of the same variables as included in the anova
m_anova <-lmer(social_cond$distance ~  Week + (1|ID), data=social_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova,pairwise ~ Week, conf.level = 0.95)
emmeans_results

class(emmeans_results)

emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
class(emm_single)

pairwise_results <- contrast(emm_single, method = "pairwise")
class(pairwise_results)

#pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")

# Perform pairwise comparisons
#pairwise_results <- contrast(emmeans_results, method = "pairwise")
# Display the results
summary(pairwise_results)

#bonferroni
#pwc <- pairs(emmeans_results, adjust = "bon")
#pwc

pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
summary(pwc)
```


```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
#significance_label <- get_significance_label(p_value)

ggplot(social_cond, aes(x = Week, y = distance)) +
  geom_beeswarm(data = social_cond, size = 2.5, aes(color = ID)) +
  ylab("Distance (cm)") +
  xlab("Timepoint social condition") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              y_position = 4600,
              annotations = c("ns")) +
  geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 4350,
              annotations = c("***")) +
  geom_signif(comparisons = list(c("1st week post induction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 4100,
              annotations = c("***")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = social_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```

#social novelty

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

novelty_cond$Week <- factor(novelty_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction"))
novelty_cond$distance = novelty_cond$`Distance moved cm`
#creating a linear model of the same variables as included in the anova
m_anova <-lmer(novelty_cond$distance ~  Week + (1|ID), data=novelty_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova,pairwise ~ Week, conf.level = 0.95)
emmeans_results

class(emmeans_results)

emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
class(emm_single)

pairwise_results <- contrast(emm_single, method = "pairwise")
class(pairwise_results)

#pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")

# Perform pairwise comparisons
#pairwise_results <- contrast(emmeans_results, method = "pairwise")
# Display the results
summary(pairwise_results)

#bonferroni
#pwc <- pairs(emmeans_results, adjust = "bon")
#pwc

pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
summary(pwc)
```

```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
#significance_label <- get_significance_label(p_value)

ggplot(novelty_cond, aes(x = Week, y = distance)) +
  geom_beeswarm(data = novelty_cond, size = 2.5, aes(color = ID)) +
  ylab("Distance (cm)") +
  xlab("Timepoint") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              y_position = 4600,
              annotations = c("ns")) +
  geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 4350,
              annotations = c("***")) +
  geom_signif(comparisons = list(c("1st week post induction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 4100,
              annotations = c("***")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = novelty_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```
