---
title: "3CT"
output:
  pdf_document: default
  html_document: default
date: '2022-12-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}

# Loading
library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans)
# xlsx files

df <- read_excel("3CT_larger_arena new.xlsx") #this file for larger arena

#df <- read_excel("3CT_larger_arena new_4weeks.xlsx")

```

Add 1cm zone with surrounding zone
```{r}
df <- df %>% mutate(social_time=(`In zone SocialZone multiCond cumulative s`+df$SocialZoneRound))

df <- df %>% mutate(novel_time=(`In zone NoveltyZone multiCond cumulative s`+df$NoveltyZoneRound))

```



```{r}

# Group assignment for each mouse_id
group_assignments <- c(
  # WT
  "224" = "WT",
  "281" = "WT",
  "262" = "WT",
  "384" = "WT",
  "439" = "WT",
  "441" = "WT",
  "443" = "WT",
  "390" = "WT",
  "392" = "WT",

  # TDP-43
  "98"  = "TDP-43",
  "102" = "TDP-43",
  "112" = "TDP-43",
  "120" = "TDP-43",
  "124" = "TDP-43",
  "128" = "TDP-43",
  "129" = "TDP-43",
  "137" = "TDP-43",
  "141" = "TDP-43",
  "247" = "TDP-43",
  "271" = "TDP-43",
  "250" = "TDP-43",
  "256" = "TDP-43",
  "382" = "TDP-43",
  "433" = "TDP-43",
  "445" = "TDP-43"
)

class(df$ID)

# Add group column to mean_per_mouse
df <- df %>%
  mutate(genotype = group_assignments[as.character(ID)])
```

# Removal of cohort 4 TG mice
```{r}

#df$ID <- as.numeric(df$ID)
#df$ID<- as.numeric(df$ID) #Change mouse number from class character to class numeric
df<-df[!(df$ID=="445"),]
df<-df[!(df$ID=="433"),]
df<-df[!(df$ID=="382"),]

#removal of other cohorts but cohort 4
# df<-df[!(df$cohort=="1"),]
# df<-df[!(df$cohort=="2"),]
# df<-df[!(df$cohort=="3"),]

```

We will primarily analyse time spent in the different zones. We expect:

Habituation phase: time spent in neutral, social, and novel zone will be equal.
For social phase (1 mouse introduced): time spent in social will be bigger than neutral and novel zone combined.
for novelty phase (1 novel mouse introduced into novelty zone): time spent in novelty zone will be bigger then in neutral and social zone combined.


Divide dataset up into Habituation, social, and novelty phases

```{r}

#use this for LMM approach with full sample size
habituation_cond <-  df %>% filter(Condition == "Hab")
social_cond<-  df %>% filter(Condition == "Social")
novelty_cond<-  df %>% filter(Condition == "Novel")
```

with 1cm zone and surrounding zone
```{r}
habituation_cond <- habituation_cond %>% mutate(social_preference_relative=((`In zone SocialZone multiCond cumulative s`-`In zone NoveltyZone multiCond cumulative s`)/(`In zone NoveltyZone multiCond cumulative s`+`In zone SocialZone multiCond cumulative s`)))



social_cond <- social_cond %>% mutate(social_preference_relative=((social_cond$social_time-social_cond$novel_time)/(social_cond$social_time+social_cond$novel_time)))

novelty_cond <- novelty_cond %>% mutate(novelty_preference_relative=((novelty_cond$novel_time-novelty_cond$social_time)/(novelty_cond$novel_time+novelty_cond$social_time)))

```


LMM sociability absolute

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

#social_cond$Week <- factor(social_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction"))
social_cond$Week <- factor(social_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

social_cond$genotype <- factor(social_cond$genotype, levels = c("WT","TDP-43"))

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(social_time ~  Week*genotype + (1|ID), data=social_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova,pairwise ~ Week*genotype, conf.level = 0.95)
emmeans_results

class(emmeans_results)

emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
class(emm_single)

pairwise_results <- contrast(emm_single, method = "pairwise")
class(pairwise_results)

#pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")

# Perform pairwise comparisons
#pairwise_results <- contrast(emmeans_results, method = "pairwise")
# Display the results
summary(pairwise_results)

#bonferroni
#pwc <- pairs(emmeans_results, adjust = "bon")
#pwc

pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
summary(pwc)

# PLANNED COMPARISONS

emm <- emmeans(m_anova, ~ Week * genotype)
# Weeks compared within each genotype
within_genotype <- pairs(emm, by = "genotype", adjust = "bon")

summary(within_genotype)

# Genotypes compared within each week
between_genotype <- pairs(emm, by = "Week", adjust = "bon")

summary(between_genotype)
```

# assumption checks
```{r}
p_load(performance)

check_model(m_anova)
```

LMM sociability pref

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

#social_cond$Week <- factor(social_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction"))


social_cond$Week <- factor(social_cond$Week, levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

social_cond$genotype <- factor(social_cond$genotype, levels = c("WT","TDP-43"))

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(social_preference_relative ~  Week*genotype + (1|ID), data=social_cond)
#m_anova <- lmer(social_preference_relative ~ Week * genotype + (1 | ID) + (1 | cohort), data = social_cond) #cohort as random effect
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova,pairwise ~ Week*genotype, conf.level = 0.95)
emmeans_results

class(emmeans_results)

emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
class(emm_single)

pairwise_results <- contrast(emm_single, method = "pairwise")
class(pairwise_results)

#pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")

# Perform pairwise comparisons
#pairwise_results <- contrast(emmeans_results, method = "pairwise")
# Display the results
summary(pairwise_results)

#bonferroni
#pwc <- pairs(emmeans_results, adjust = "bon")
#pwc

pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
summary(pwc)
# PLANNED COMPARISONS

emm <- emmeans(m_anova, ~ Week * genotype)
# Weeks compared within each genotype
within_genotype <- pairs(emm, by = "genotype", adjust = "bon")

summary(within_genotype)

# Genotypes compared within each week
between_genotype <- pairs(emm, by = "Week", adjust = "bon")

summary(between_genotype)

# PLANNED COMPARISONS

emm <- emmeans(m_anova, ~ Week * genotype)
# Weeks compared within each genotype
within_genotype <- pairs(emm, by = "genotype", adjust = "bon")

summary(within_genotype)

#store the p values

# Genotypes compared within each week
between_genotype <- pairs(emm, by = "Week", adjust = "bon")

summary(between_genotype)
```

#store the p values
```{r}

# Get the summary of the within_genotype object
within_sum <- summary(within_genotype)

# View to understand structure
within_sum
# It should show a data.frame/tbl with columns: contrast, estimate, SE, df, t.ratio, p.value

# ----------------------------
# 1️⃣ Extract p-values for TDP comparisons
# ----------------------------
TDP_pre_1 <- within_sum$p.value[within_sum$genotype == "TDP-43" & within_sum$contrast == "Preinduction - 1st week post induction"]
TDP_pre_2 <- within_sum$p.value[within_sum$genotype == "TDP-43" & within_sum$contrast == "Preinduction - 2nd week post induction"]
TDP_1_2   <- within_sum$p.value[within_sum$genotype == "TDP-43" & within_sum$contrast == "1st week post induction - 2nd week post induction"]

# ----------------------------
# 2️⃣ Extract p-values for WT comparisons (optional)
# ----------------------------
WT_pre_1 <- within_sum$p.value[within_sum$genotype == "WT" & within_sum$contrast == "Preinduction - 1st week post induction"]
WT_pre_2 <- within_sum$p.value[within_sum$genotype == "WT" & within_sum$contrast == "Preinduction - 2nd week post induction"]
WT_1_2   <- within_sum$p.value[within_sum$genotype == "WT" & within_sum$contrast == "1st week post induction - 2nd week post induction"]

# ----------------------------
# 3️⃣ Check
# ----------------------------
TDP_pre_1
TDP_pre_2
TDP_1_2

WT_pre_1
WT_pre_2
WT_1_2


```

```{r}

levels(df$Week)[levels(df$Week) == "1st week post induction"] <- "1st week"
levels(df$Week)[levels(df$Week) == "2nd week post induction"] <- "2nd week"
levels(df$Week)[levels(df$Week) == "3rd week post induction"] <- "3rd week"

# ----------------------------
# 1️⃣ Get summary of between_genotype object
# ----------------------------
between_sum <- summary(between_genotype)

# View to understand structure
between_sum
# It should show columns: contrast, estimate, SE, df, t.ratio, p.value, Week

# ----------------------------
# 2️⃣ Extract p-values for each week comparison between WT and TDP
# ----------------------------

# Preinduction: WT vs TDP
Pre_vs_TDP_pre <- between_sum$p.value[between_sum$Week == "Preinduction" & between_sum$contrast == "WT - (TDP-43)"]

# 1st week: WT vs TDP
Pre_vs_TDP_w1 <- between_sum$p.value[between_sum$Week == "1st week post induction" & between_sum$contrast == "WT - (TDP-43)"]

# 2nd week: WT vs TDP
Pre_vs_TDP_w2 <- between_sum$p.value[between_sum$Week == "2nd week post induction" & between_sum$contrast == "WT - (TDP-43)"]

# ----------------------------
# 3️⃣ Check
# ----------------------------
Pre_vs_TDP_pre
Pre_vs_TDP_w1
Pre_vs_TDP_w2

```




# assumption checks
```{r}
p_load(performance)

check <- check_model(m_anova)
check
```

#assumption checks
```{r}
plot(fitted(m_anova), residuals(m_anova))  # Should look like random scatter



qqnorm(residuals(m_anova)); qqline(residuals(m_anova))  # Q-Q plot - some tails
shapiro.test(residuals(m_anova))  # Shapiro-Wilk test (for small samples)

```
plot

```{r}
# cbbPalette <- c("#fcfdbf", "#feca8d", "#D67236", "#f1605d", "#9e2f7f", "#440f76")
# 
# p_load(ggbeeswarm, ggsignif)
# levels(social_cond$Week)[levels(social_cond$Week) == "1st week post induction"] <- "1st week"
# levels(social_cond$Week)[levels(social_cond$Week) == "2nd week post induction"] <- "2nd week"
# levels(social_cond$Week)[levels(social_cond$Week) == "3rd week post induction"] <- "3rd week"
# 
# #ggplot(social_cond, aes(x = Week, y = social_time)) +
# ggplot(social_cond, aes(x = Week, y = social_preference_relative)) +
#   geom_beeswarm(data = social_cond, size = 2.5, aes(color = ID)) +
#   ylab("Social Preference") +
#   xlab("Timepoint") +
#   # geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
#   #             map_signif_level = TRUE,
#   #             y_position = 1.3,
#   #             annotations = c("0.1283")) +
#   # geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
#   #             map_signif_level = TRUE,
#   #             y_position = 1.15,
#   #             annotations = c(" 0.0087")) +
#   # geom_signif(comparisons = list(c("1st week post induction", "2nd week post induction")),
#   #             map_signif_level = TRUE,
#   #             y_position = 1,
#   #             annotations = c("0.4487")) +
# facet_wrap(~genotype)+
#   geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
#   geom_errorbar(data = social_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
#   theme(text = element_text(size = 13))


```

```{r}
library(ggplot2)
library(ggdist)
library(ggsignif)
library(dplyr)

levels(social_cond$Week)[levels(social_cond$Week) == "1st week post induction"] <- "1st week"
levels(social_cond$Week)[levels(social_cond$Week) == "2nd week post induction"] <- "2nd week"
levels(social_cond$Week)[levels(social_cond$Week) == "3rd week post induction"] <- "3rd week"

# ----------------------------
# 1️⃣ Combined x-axis variable
# ----------------------------
social_cond$Week_genotype <- interaction(social_cond$genotype, social_cond$Week, sep = "_", lex.order = TRUE)

# X-axis labels (weeks only)
week_labels <- c("Preinduction", "1st week", "2nd week")
x_labels <- rep(week_labels, times = length(levels(social_cond$genotype)))

# ----------------------------
# 2️⃣ Define comparisons
# ----------------------------

# Within-genotype comparisons (already done)
comparisons_within <- list(
  c("WT_Preinduction", "WT_1st week"),
  c("WT_1st week", "WT_2nd week"),
  c("WT_Preinduction", "WT_2nd week"),
  c("TDP-43_Preinduction", "TDP-43_1st week"),
  c("TDP-43_1st week", "TDP-43_2nd week"),
  c("TDP-43_Preinduction", "TDP-43_2nd week")
)

# Between-genotype comparisons (WT vs TDP at each week)
comparisons_between <- list(
  c("WT_Preinduction", "TDP-43_Preinduction"),
  c("WT_1st week", "TDP-43_1st week"),
  c("WT_2nd week", "TDP-43_2nd week")
)

# ----------------------------
# 3️⃣ Compute staggered y_positions
# ----------------------------

y_max_WT  <- max(social_cond$social_preference_relative[social_cond$genotype == "WT"], na.rm = TRUE)
y_max_TDP <- max(social_cond$social_preference_relative[social_cond$genotype == "TDP-43"], na.rm = TRUE)
y_max_all <- max(social_cond$social_preference_relative, na.rm = TRUE)  # for between-genotype

# Within-genotype: stagger 5%,10%,15% above each genotype's max
y_positions_within <- c(
  y_max_WT  * (1 + c(0.05, 0.10, 0.15)),
  y_max_TDP * (1 + c(0.05, 0.10, 0.15))
)

# Between-genotype: stagger 5%,10%,15% above overall max
y_positions_between <- y_max_all * (1 + c(0.20, 0.25, 0.30))

# ----------------------------
# 4️⃣ Combine into one data.frame for geom_signif with P= and 4 decimals
# ----------------------------

sig_social_cond <- data.frame(
  xmin = c(sapply(comparisons_within, `[`, 1), sapply(comparisons_between, `[`, 1)),
  xmax = c(sapply(comparisons_within, `[`, 2), sapply(comparisons_between, `[`, 2)),
  y_position = c(y_positions_within, y_positions_between),
  annotations = c(
    # within-genotype p-values with "P =" and 4 decimals
    paste0("P = ", formatC(c(WT_pre_1, WT_1_2, WT_pre_2, TDP_pre_1, TDP_1_2, TDP_pre_2),
                            format = "f", digits = 4)),
    # between-genotype p-values with "P =" and 4 decimals
    paste0("P = ", formatC(c(Pre_vs_TDP_pre, Pre_vs_TDP_w1, Pre_vs_TDP_w2),
                            format = "f", digits = 4))
  )
)

# sig_social_cond <- data.frame(
#   xmin = c(sapply(comparisons_within, `[`, 1), sapply(comparisons_between, `[`, 1)),
#   xmax = c(sapply(comparisons_within, `[`, 2), sapply(comparisons_between, `[`, 2)),
#   y_position = c(y_positions_within, y_positions_between),
#   annotations = c(
#     # within-genotype p-values
#     format(round(c(WT_pre_1, WT_1_2, WT_pre_2, TDP_pre_1, TDP_1_2, TDP_pre_2), 2), nsmall = 2),
#     # between-genotype p-values
#     format(round(c(Pre_vs_TDP_pre, Pre_vs_TDP_w1, Pre_vs_TDP_w2), 2), nsmall = 2)
#   )
# )

# ----------------------------
# 5️⃣ Raincloud plot with all significance bars
# ----------------------------

ggplot(social_cond, aes(x = Week_genotype, y = social_preference_relative, color = genotype, fill = genotype)) +

  # half violin
  ggdist::stat_halfeye(
    side = "left",
    adjust = 0.5,
    justification = 1.1,
    .width = 0,
    point_colour = NA,
    alpha = 0.6,
    scale = 0.5
  ) +

  # connect repeated IDs
  geom_line(aes(group = ID), color = "gray50", linewidth = 0.5, alpha = 0.4) +

  # individual datapoints
  geom_jitter(width = 0.05, size = 2.5, alpha = 0.6) +

  # mean ± SE
  geom_crossbar(stat = "summary",
                fun = mean,
                width = 0.2,
                fatten = 1.5,
                color = "black",
                fill = "black") +

  geom_errorbar(stat = "summary",
                fun.data = mean_se,
                width = 0.1,
                color = "black",
                fill = "black") +

  # significance bars using manual mode
  geom_signif(
    data = sig_social_cond,
    aes(xmin = xmin, xmax = xmax, annotations = annotations, y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 4,
    inherit.aes = FALSE
  ) +

  # replace x-axis labels
  scale_x_discrete(labels = x_labels) +

  # labels and theme
  ylab("sociability Preference") +
  xlab("Timepoint") +
  theme(text = element_text(size = 16))


# ----------------------------
# 6️⃣ Save the plot in vector format
# ----------------------------

# Option 1: Save as PDF
ggsave(
  filename = "raincloud_plot_3CT_social.pdf",  # filename
  plot = last_plot(),               # current last plot
  width = 10,                       # width in inches
  height = 6,                       # height in inches
  device = "pdf"                    # vector format
)

# Option 2: Save as SVG
ggsave(
  filename = "raincloud_plot_3CT_social.svg",
  plot = last_plot(),
  width = 10,
  height = 6,
  device = "svg"
)

```


LLM absolute novel

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

novelty_cond$genotype <- factor(novelty_cond$genotype, levels = c("WT","TDP-43"))

#novelty_cond$Week <- factor(novelty_cond$Week , levels = c("Preinduction","1st week post induction","2nd week #post induction"))
novelty_cond$Week <- factor(novelty_cond$Week , levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(novel_time ~  Week*genotype + (1|ID), data=novelty_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week*genotype)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc

# PLANNED COMPARISONS

emm <- emmeans(m_anova, ~ Week * genotype)
# Weeks compared within each genotype
within_genotype <- pairs(emm, by = "genotype", adjust = "bon")

summary(within_genotype)

# Genotypes compared within each week
between_genotype <- pairs(emm, by = "Week", adjust = "bon")

summary(between_genotype)
```

# assumption checks
```{r}
p_load(performance)

check_model(m_anova)
```

LLM novelty pref

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

#novelty_cond$Week <- factor(novelty_cond$Week , levels = c("Preinduction","1st week post induction","2nd week #post induction"))

novelty_cond$Week <- factor(novelty_cond$Week , levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(novelty_preference_relative ~  Week*genotype + (1|ID), data=novelty_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week*genotype)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc

# PLANNED COMPARISONS

emm <- emmeans(m_anova, ~ Week * genotype)
# Weeks compared within each genotype
within_genotype <- pairs(emm, by = "genotype", adjust = "bon")

summary(within_genotype)

# Genotypes compared within each week
between_genotype <- pairs(emm, by = "Week", adjust = "bon")

summary(between_genotype)
```


#store the p values
```{r}

# Get the summary of the within_genotype object
within_sum <- summary(within_genotype)

# View to understand structure
within_sum
# It should show a data.frame/tbl with columns: contrast, estimate, SE, df, t.ratio, p.value

# ----------------------------
# 1️⃣ Extract p-values for TDP comparisons
# ----------------------------
TDP_pre_1 <- within_sum$p.value[within_sum$genotype == "TDP-43" & within_sum$contrast == "Preinduction - 1st week post induction"]
TDP_pre_2 <- within_sum$p.value[within_sum$genotype == "TDP-43" & within_sum$contrast == "Preinduction - 2nd week post induction"]
TDP_1_2   <- within_sum$p.value[within_sum$genotype == "TDP-43" & within_sum$contrast == "1st week post induction - 2nd week post induction"]

# ----------------------------
# 2️⃣ Extract p-values for WT comparisons (optional)
# ----------------------------
WT_pre_1 <- within_sum$p.value[within_sum$genotype == "WT" & within_sum$contrast == "Preinduction - 1st week post induction"]
WT_pre_2 <- within_sum$p.value[within_sum$genotype == "WT" & within_sum$contrast == "Preinduction - 2nd week post induction"]
WT_1_2   <- within_sum$p.value[within_sum$genotype == "WT" & within_sum$contrast == "1st week post induction - 2nd week post induction"]

# ----------------------------
# 3️⃣ Check
# ----------------------------
TDP_pre_1
TDP_pre_2
TDP_1_2

WT_pre_1
WT_pre_2
WT_1_2


```

```{r}

levels(df$Week)[levels(df$Week) == "1st week post induction"] <- "1st week"
levels(df$Week)[levels(df$Week) == "2nd week post induction"] <- "2nd week"
levels(df$Week)[levels(df$Week) == "3rd week post induction"] <- "3rd week"

# ----------------------------
# 1️⃣ Get summary of between_genotype object
# ----------------------------
between_sum <- summary(between_genotype)

# View to understand structure
between_sum
# It should show columns: contrast, estimate, SE, df, t.ratio, p.value, Week

# ----------------------------
# 2️⃣ Extract p-values for each week comparison between WT and TDP
# ----------------------------

# Preinduction: WT vs TDP
Pre_vs_TDP_pre <- between_sum$p.value[between_sum$Week == "Preinduction" & between_sum$contrast == "WT - (TDP-43)"]

# 1st week: WT vs TDP
Pre_vs_TDP_w1 <- between_sum$p.value[between_sum$Week == "1st week post induction" & between_sum$contrast == "WT - (TDP-43)"]

# 2nd week: WT vs TDP
Pre_vs_TDP_w2 <- between_sum$p.value[between_sum$Week == "2nd week post induction" & between_sum$contrast == "WT - (TDP-43)"]

# ----------------------------
# 3️⃣ Check
# ----------------------------
Pre_vs_TDP_pre
Pre_vs_TDP_w1
Pre_vs_TDP_w2

```



# assumption checks
```{r}
p_load(performance)

check <- check_model(m_anova)
check
```

#assumption checks
```{r}
plot(fitted(m_anova), residuals(m_anova))  # Should look like random scatter



qqnorm(residuals(m_anova)); qqline(residuals(m_anova))  # Q-Q plot - some tails
shapiro.test(residuals(m_anova))  # Shapiro-Wilk test (for small samples)

```


```{r}
library(ggplot2)
library(ggdist)
library(ggsignif)
library(dplyr)

levels(novelty_cond$Week)[levels(novelty_cond$Week) == "1st week post induction"] <- "1st week"
levels(novelty_cond$Week)[levels(novelty_cond$Week) == "2nd week post induction"] <- "2nd week"
levels(novelty_cond$Week)[levels(novelty_cond$Week) == "3rd week post induction"] <- "3rd week"

# ----------------------------
# 1️⃣ Combined x-axis variable
# ----------------------------
novelty_cond$Week_genotype <- interaction(novelty_cond$genotype, novelty_cond$Week, sep = "_", lex.order = TRUE)

# X-axis labels (weeks only)
week_labels <- c("Preinduction", "1st week", "2nd week")
x_labels <- rep(week_labels, times = length(levels(novelty_cond$genotype)))

# ----------------------------
# 2️⃣ Define comparisons
# ----------------------------

# Within-genotype comparisons (already done)
comparisons_within <- list(
  c("WT_Preinduction", "WT_1st week"),
  c("WT_1st week", "WT_2nd week"),
  c("WT_Preinduction", "WT_2nd week"),
  c("TDP-43_Preinduction", "TDP-43_1st week"),
  c("TDP-43_1st week", "TDP-43_2nd week"),
  c("TDP-43_Preinduction", "TDP-43_2nd week")
)

# Between-genotype comparisons (WT vs TDP at each week)
comparisons_between <- list(
  c("WT_Preinduction", "TDP-43_Preinduction"),
  c("WT_1st week", "TDP-43_1st week"),
  c("WT_2nd week", "TDP-43_2nd week")
)

# ----------------------------
# 3️⃣ Compute staggered y_positions
# ----------------------------

y_max_WT  <- max(novelty_cond$novelty_preference_relative[novelty_cond$genotype == "WT"], na.rm = TRUE)
y_max_TDP <- max(novelty_cond$novelty_preference_relative[novelty_cond$genotype == "TDP-43"], na.rm = TRUE)
y_max_all <- max(novelty_cond$novelty_preference_relative, na.rm = TRUE)  # for between-genotype

# Within-genotype: stagger 5%,10%,15% above each genotype's max
y_positions_within <- c(
  y_max_WT  * (1 + c(0.05, 0.10, 0.15)),
  y_max_TDP * (1 + c(0.05, 0.10, 0.15))
)

# Between-genotype: stagger 5%,10%,15% above overall max
y_positions_between <- y_max_all * (1 + c(0.20, 0.25, 0.30))

# ----------------------------
# 4️⃣ Combine into one data.frame for geom_signif with P= and 4 decimals
# ----------------------------

sig_novelty_cond <- data.frame(
  xmin = c(sapply(comparisons_within, `[`, 1), sapply(comparisons_between, `[`, 1)),
  xmax = c(sapply(comparisons_within, `[`, 2), sapply(comparisons_between, `[`, 2)),
  y_position = c(y_positions_within, y_positions_between),
  annotations = c(
    # within-genotype p-values with "P =" and 4 decimals
    paste0("P = ", formatC(c(WT_pre_1, WT_1_2, WT_pre_2, TDP_pre_1, TDP_1_2, TDP_pre_2),
                            format = "f", digits = 4)),
    # between-genotype p-values with "P =" and 4 decimals
    paste0("P = ", formatC(c(Pre_vs_TDP_pre, Pre_vs_TDP_w1, Pre_vs_TDP_w2),
                            format = "f", digits = 4))
  )
)

# sig_novelty_cond <- data.frame(
#   xmin = c(sapply(comparisons_within, `[`, 1), sapply(comparisons_between, `[`, 1)),
#   xmax = c(sapply(comparisons_within, `[`, 2), sapply(comparisons_between, `[`, 2)),
#   y_position = c(y_positions_within, y_positions_between),
#   annotations = c(
#     # within-genotype p-values
#     format(round(c(WT_pre_1, WT_1_2, WT_pre_2, TDP_pre_1, TDP_1_2, TDP_pre_2), 2), nsmall = 2),
#     # between-genotype p-values
#     format(round(c(Pre_vs_TDP_pre, Pre_vs_TDP_w1, Pre_vs_TDP_w2), 2), nsmall = 2)
#   )
# )

# ----------------------------
# 5️⃣ Raincloud plot with all significance bars
# ----------------------------

ggplot(novelty_cond, aes(x = Week_genotype, y = novelty_preference_relative, color = genotype, fill = genotype)) +

  # half violin
  ggdist::stat_halfeye(
    side = "left",
    adjust = 0.5,
    justification = 1.1,
    .width = 0,
    point_colour = NA,
    alpha = 0.6,
    scale = 0.5
  ) +

  # connect repeated IDs
  geom_line(aes(group = ID), color = "gray50", linewidth = 0.5, alpha = 0.4) +

  # individual datapoints
  geom_jitter(width = 0.05, size = 2.5, alpha = 0.6) +

  # mean ± SE
  geom_crossbar(stat = "summary",
                fun = mean,
                width = 0.2,
                fatten = 1.5,
                color = "black",
                fill = "black") +

  geom_errorbar(stat = "summary",
                fun.data = mean_se,
                width = 0.1,
                color = "black",
                fill = "black") +

  # significance bars using manual mode
  geom_signif(
    data = sig_novelty_cond,
    aes(xmin = xmin, xmax = xmax, annotations = annotations, y_position = y_position),
    manual = TRUE,
    tip_length = 0.02,
    textsize = 4,
    inherit.aes = FALSE
  ) +

  # replace x-axis labels
  scale_x_discrete(labels = x_labels) +

  # labels and theme
  ylab("social novelty preference") +
  xlab("Timepoint") +
  theme(text = element_text(size = 16))


# ----------------------------
# 6️⃣ Save the plot in vector format
# ----------------------------

# Option 1: Save as PDF
ggsave(
  filename = "raincloud_plot_3CT_novel.pdf",  # filename
  plot = last_plot(),               # current last plot
  width = 10,                       # width in inches
  height = 6,                       # height in inches
  device = "pdf"                    # vector format
)

# Option 2: Save as SVG
ggsave(
  filename = "raincloud_plot_3CT_novel.svg",
  plot = last_plot(),
  width = 10,
  height = 6,
  device = "svg"
)

```

plot

```{r}
# cbbPalette <- c("#fcfdbf", "#feca8d", "#D67236", "#f1605d", "#9e2f7f", "#440f76")
# 
# p_load(ggbeeswarm, ggsignif)
# levels(novelty_cond$Week)[levels(novelty_cond$Week) == "1st week post induction"] <- "1st week"
# levels(novelty_cond$Week)[levels(novelty_cond$Week) == "2nd week post induction"] <- "2nd week"
# levels(novelty_cond$Week)[levels(novelty_cond$Week) == "3rd week post induction"] <- "3rd week"
# 
# ggplot(novelty_cond, aes(x = Week, y = novelty_preference_relative)) +
#   geom_beeswarm(data = novelty_cond, size = 2.5, aes(color = ID)) +
#   ylab("Novelty Preference") +
#   xlab("Timepoint") +
#   # geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
#   #             map_signif_level = TRUE,
#   #             y_position = 1.3,
#   #             annotations = c("0.7010")) +
#   # geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
#   #             map_signif_level = TRUE,
#   #             y_position = 1.15,
#   #             annotations = c("0.2114")) +
#   # geom_signif(comparisons = list(c("1st week post induction", "2nd week post induction")),
#   #             map_signif_level = TRUE,
#   #             y_position = 1,
#   #             annotations = c("1.0000")) +
# facet_wrap(~genotype) +
#   geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
#   geom_errorbar(data = novelty_cond, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
#   theme(text = element_text(size = 13))

```



# COHORT COMPARISON

```{r}

library(ggplot2)
library(ggbeeswarm)
library(dplyr)


# Plot: Dark preference by Week, colored by genotype, faceted by cohort
ggplot(social_cond, aes(x = Week, y = social_cond$social_preference_relative, color = genotype)) +
  geom_beeswarm(size = 2.5, cex = 1.1, priority = "density") +
  stat_summary(
    aes(group = Week),
    fun = mean,
    geom = "crossbar",
    color = "black",
    width = 0.25,
    fatten = 1.3,
    show.legend = FALSE
  ) +
  stat_summary(
    aes(group = Week),
    fun.data = mean_se,
    geom = "errorbar",
    color = "black",
    width = 0.12,
    show.legend = FALSE
  ) +
  facet_wrap(~ cohort) +
  scale_y_continuous(
    limits = c(-1, 1),
    expand = expansion(mult = c(0.02, 0.06))
  ) +
  labs(
    x = "Timepoint",
    y = "Social preference",
    color = "Genotype",
    title = "Social preference across weeks by cohort"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "bold")
  )

```



```{r}

library(ggplot2)
library(ggbeeswarm)
library(dplyr)


novelty_cond$genotype <- factor(novelty_cond$genotype, levels = c("TDP-43", "WT"))

# Plot: Dark preference by Week, colored by genotype, faceted by cohort
ggplot(novelty_cond, aes(x = Week, y = novelty_cond$novelty_preference_relative, color = genotype)) +
  geom_beeswarm(size = 2.5, cex = 1.1, priority = "density") +
  stat_summary(
    aes(group = Week),
    fun = mean,
    geom = "crossbar",
    color = "black",
    width = 0.25,
    fatten = 1.3,
    show.legend = FALSE
  ) +
  stat_summary(
    aes(group = Week),
    fun.data = mean_se,
    geom = "errorbar",
    color = "black",
    width = 0.12,
    show.legend = FALSE
  ) +
  facet_wrap(~ cohort) +
  scale_y_continuous(
    limits = c(-1, 1),
    expand = expansion(mult = c(0.02, 0.06))
  ) +
  labs(
    x = "Timepoint",
    y = "Novel preference",
    color = "Genotype",
    title = "Novel preference across weeks by cohort"
  ) +
  theme_minimal(base_size = 13) +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "right",
    strip.text = element_text(face = "bold")
  )

```


# SUBTRACTION METHOD

novelty_preference_relative

distance
```{r}
soc_norm <- social_cond %>%
  group_by(ID) %>%
  mutate(social_preference_relative = social_preference_relative - social_preference_relative[Week == "Preinduction"]) %>%
  ungroup() %>%
  filter(Week %in% c("1st week", "2nd week"))

```


#LMM of subtracted data
Sociability
```{r}

# Mixed model (correct for repeated measures)
m_anova <- lmer(social_preference_relative ~ Week * genotype + (1 | ID), data = soc_norm)

# Estimated marginal means for genotype within each week
emm <- emmeans(m_anova, ~ genotype | Week)

# WT vs TDP-43 comparisons for each week (repeated-measures aware)
genotype_comparisons <- contrast(emm, method = "pairwise", adjust = "tukey")

summary(genotype_comparisons)


```

#plot
```{r}
library(ggplot2)
library(ggbeeswarm)
library(dplyr)

ggplot(soc_norm, aes(x = Week, y = social_preference_relative, color = ID)) +
  facet_wrap(~ genotype) +
  geom_beeswarm(size = 2.5, cex = 1.2, alpha = 0.9) +

  # Mean crossbar
  geom_crossbar(
    stat = "summary",
    fun = mean,
    width = 0.2,
    fatten = 1.5,
    color = "black"
  ) +

  # Mean ± SE error bars
  geom_errorbar(
    stat = "summary",
    fun.data = mean_se,
    width = 0.1,
    color = "black"
  ) +

  labs(
    x = "Timepoint",
    y = "novelty_preference_relative"
  ) +

  theme_bw() +
  theme(
    text = element_text(size = 16),
    strip.text = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14),
    panel.grid = element_blank(),
    legend.position = "none"   # remove ID color legend (cleaner)
  )

```

# SUBTRACTION METHOD




```{r}
nov_norm <- novelty_cond %>%
  group_by(ID) %>%
  mutate(novelty_preference_relative = novelty_preference_relative - novelty_preference_relative[Week == "Preinduction"]) %>%
  ungroup() %>%
  filter(Week %in% c("1st week", "2nd week"))

```


#LMM of subtracted data
Social novelty
```{r}

# Mixed model (correct for repeated measures)
m_anova <- lmer(novelty_preference_relative ~ Week * genotype + (1 | ID), data = nov_norm)

# Estimated marginal means for genotype within each week
emm <- emmeans(m_anova, ~ genotype | Week)

# WT vs TDP-43 comparisons for each week (repeated-measures aware)
genotype_comparisons <- contrast(emm, method = "pairwise", adjust = "tukey")

summary(genotype_comparisons)


```
#plot
```{r}
library(ggplot2)
library(ggbeeswarm)
library(dplyr)

ggplot(nov_norm, aes(x = Week, y = novelty_preference_relative, color = ID)) +
  facet_wrap(~ genotype) +
  geom_beeswarm(size = 2.5, cex = 1.2, alpha = 0.9) +

  # Mean crossbar
  geom_crossbar(
    stat = "summary",
    fun = mean,
    width = 0.2,
    fatten = 1.5,
    color = "black"
  ) +

  # Mean ± SE error bars
  geom_errorbar(
    stat = "summary",
    fun.data = mean_se,
    width = 0.1,
    color = "black"
  ) +

  labs(
    x = "Timepoint",
    y = "novelty_preference_relative"
  ) +

  theme_bw() +
  theme(
    text = element_text(size = 16),
    strip.text = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 18),
    axis.text = element_text(size = 14),
    panel.grid = element_blank(),
    legend.position = "none"   # remove ID color legend (cleaner)
  )

```