---
title: "3CT"
output:
  pdf_document: default
  html_document: default
date: '2022-12-02'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load data
```{r}

# Loading
library(pacman)
p_load(readxl, tidyverse)
p_load(broom, ggsignif, multcomp, emmeans)

df <- read_excel("3CT_larger_arena_FP.xlsx") #this fiel for larger arena

```

Add 1cm zone with surrounding zone
```{r}
df <- df %>% mutate(social_time=(`In zone SocialZone multiCond cumulative s`+df$SocialZoneRound))

df <- df %>% mutate(novel_time=(`In zone NoveltyZone multiCond cumulative s`+df$NoveltyZoneRound))

```


Removal of mouse 224 from dataset it turned it it was not double transgenic
```{r}
# remove WT mice
df<-df[!(df$ID==429),]
df<-df[!(df$ID==374),]
df<-df[!(df$ID==461),]
df<-df[!(df$ID==463),]
df<-df[!(df$ID==402),]


#sorting away PFC animals

df<-df[!(df$ID==423),]
df<-df[!(df$ID==376),]
df<-df[!(df$ID==372),]
df<-df[!(df$ID==398),]
df<-df[!(df$ID==400),]
df<-df[!(df$ID==459),]
df<-df[!(df$ID==450),]
df<-df[!(df$ID==584),]
df<-df[!(df$ID==597),]


#sorting away insula animals
# df<-df[!(df$ID==533),]
# df<-df[!(df$ID==541),]
# df<-df[!(df$ID==580),]
# df<-df[!(df$ID==566),]
# df<-df[!(df$ID==588),]
# df<-df[!(df$ID==590),]

```


We will primarily analyse time spent in the different zones. We expect:

Habituation phase: time spent in neutral, social, and novel zone will be equal.
For social phase (1 mouse introduced): time spent in social will be bigger than neutral and novel zone combined.
for novelty phase (1 novel mouse introduced into novelty zone): time spent in novelty zone will be bigger then in neutral and social zone combined.


Divide dataset up into Habituation, social, and novelty phases

```{r}

#use this for LMM approach with full sample size
habituation_cond <-  df %>% filter(Condition == "Hab")
social_cond<-  df %>% filter(Condition == "Social")
novelty_cond<-  df %>% filter(Condition == "Novel")
```

with 1cm zone and surrounding zone
```{r}
habituation_cond <- habituation_cond %>% mutate(social_preference_relative=((`In zone SocialZone multiCond cumulative s`-`In zone NoveltyZone multiCond cumulative s`)/(`In zone NoveltyZone multiCond cumulative s`+`In zone SocialZone multiCond cumulative s`)))



social_cond <- social_cond %>% mutate(social_preference_relative=((social_cond$social_time-social_cond$novel_time)/(social_cond$social_time+social_cond$novel_time)))

novelty_cond <- novelty_cond %>% mutate(novelty_preference_relative=((novelty_cond$novel_time-novelty_cond$social_time)/(novelty_cond$novel_time+novelty_cond$social_time)))

```

LMM sociability

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

social_cond$Week <- factor(social_cond$Week , levels = c("Preinduction","1st week post induction","2nd week post induction"))

levels(social_cond$Week)[levels(social_cond$Week) == "1st week post induction"] <- "1st week"
levels(social_cond$Week)[levels(social_cond$Week) == "2nd week post induction"] <- "2nd week"


#creating a linear model of the same variables as included in the anova
m_anova <-lmer(social_preference_relative ~  Week + (1|ID), data=social_cond)

summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week)

# for plot
emmeans_results_df <- as.data.frame(emmeans_results)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc

# # for plot
# emmeans_results_df <- as.data.frame(emmeans_results)
# 
# emm_single <- emmeans_results[[1]]  # Extract the first emmGrid object
# class(emm_single)
# 
# pairwise_results <- contrast(emm_single, method = "pairwise")
# class(pairwise_results)
# 
# #pairwise_results <- lapply(emmeans_results, contrast, method = "pairwise")
# 
# # Perform pairwise comparisons
# #pairwise_results <- contrast(emmeans_results, method = "pairwise")
# # Display the results
# summary(pairwise_results)
# 
# #bonferroni
# #pwc <- pairs(emmeans_results, adjust = "bon")
# #pwc
# 
# pwc <- contrast(emm_single, method = "pairwise", adjust = "bon")
# summary(pwc)
```

 


# assumption checks
```{r}
p_load(performance)

check <- check_model(m_anova)
check
```

#assumption checks
```{r}

qqnorm(residuals(m_anova)); qqline(residuals(m_anova))  # Q-Q plot - some tails
shapiro.test(residuals(m_anova))  # Shapiro-Wilk test (for small samples)

```


#assumption checks


```{r}
#assumption check 1 - Normality of residuals (should be normally distributed) - violates
qqnorm(resid(m_anova))
qqline(resid(m_anova))
shapiro.test(residuals(m_anova)) 
#probs a lil' bit non-normal, thus, when calculating confidence intervals, used "boot" method which assumes non-normality
```

```{r}
#checks for distribution, compliments residual plot
hist(resid(m_anova)) 
```

```{r}
#Assumption 2 - Homoscedasticity (variance should be consistent) - passes
plot(fitted(m_anova), resid(m_anova))
#meets assumption, evenly distributed, thus, Homoscedasticity
```

```{r}
#When you plot residuals by ID, you're looking for:
#No systematic pattern across individuals.
#No obvious trends within IDs (e.g., all high or all low residuals).
#Ideally, residuals centered around zero with some variation.
boxplot(resid(m_anova) ~ social_cond$ID)
```

```{r}
#Assumption 3 - confidence intervals - passes
confint(m_anova, method = "boot")
#if CI includes zero between 2.5% and 97.5% then that effect is not statistically significant.

```


# Raincloud

```{r}
# Convert emmeans pairwise comparisons to data frame (BONFERRONI already applied in pwc)
pairwise_results_oft <- as.data.frame(pwc)

# Filter significant results (optional, not required for plotting)
sig_results_oft <- pairwise_results_oft %>%
  dplyr::filter(p.value < 0.05)

# Add significance labels
sig_labels_oft <- pairwise_results_oft %>%
  dplyr::mutate(
    label = dplyr::case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

# View labeled results
sig_labels_oft
```



```{r}

library(dplyr)
library(tidyr)

# IMPORTANT: ensure this is a data.frame
pairwise_results_oft <- as.data.frame(pwc)

# pre-compute safe y-position reference (NEVER rely on df inside pipe)
y_max <- max(social_cond$social_preference_relative, na.rm = TRUE)

sig_labels_oft <- pairwise_results_oft %>%
  
  # split contrasts into groups
  separate(contrast, into = c("group1", "group2"), sep = " - ") %>%
  
  mutate(
    # significance labels (Bonferroni already in pwc)
    label = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    
    # safe, automatic y positioning
    y.position = y_max * (1 + 0.10 * row_number())
  ) %>%
  
  # explicitly use dplyr::select to avoid masking issues
  dplyr::select(group1, group2, y.position, label)

sig_labels_oft

```
```{r}
emmeans_results_df
```

```{r}
p_load(ggpubr, ggdist)

plot_raincloud <- ggplot(social_cond, aes(x = Week, y = social_preference_relative, fill = Week, color = Week)) +
  
  # Half-violin, nudged left
  stat_halfeye(
    side = "left",
    adjust = 0.5,
    justification = 1.1,
    .width = 0,
    point_colour = NA,
    alpha = 0.6,
    scale = 0.5, #make plot narrower
    position = position_nudge(x = -0.05)
  ) +
  
#  # Boxplot
#  geom_boxplot(
#    width = 0.05,
#    outlier.shape = NA,
#    alpha = 0.7,
#    color = "black",
#    show.legend=FALSE,
#    position = position_nudge(x = -0.15)
#  ) +
  
  
  #connect IDs
  geom_line(
  aes(x = Week, y = social_preference_relative, group = ID),
  color = "gray50",
  linewidth = 0.5,
  alpha = 0.4
) +

 # Jittered points
  geom_jitter(
    width = 0.04,
    size = 2.0,
    alpha = 0.5,
  ) +
  
  #add means for SE bars
  geom_point(
    data = emmeans_results_df,
    aes(x = Week, y = emmean),
    color = "black",
    size = 3.0,
    inherit.aes = FALSE,
    position = position_nudge(x = -0.15)
  ) +
  
    
  #adding in model SE data
  geom_errorbar(
    data = emmeans_results_df,
    aes(x = Week, ymin = emmean - SE, ymax = emmean + SE),
    width = 0.10,
    inherit.aes = FALSE,
    linewidth = 0.6,
    position = position_nudge(x = -0.15)
  ) +
  
  
  # # Add in significant lines and asterisks
  # stat_pvalue_manual(
  #   sig_labels_oft,
  #   label = "p.adj.signif",
  #   tip.length = 0.02,
  #   size = 6,
  #   inherit.aes = FALSE
  # ) +
  
  stat_pvalue_manual(
    sig_labels_oft,
    label = "label",
    tip.length = 0.02,
    size = 6,
    inherit.aes = FALSE
  ) +

  # Manual colors
  scale_fill_manual(values = c("Preinduction" = "#feca8d", "1st week" = "#f1605d", "2nd week" = "#440f76")) +
  scale_color_manual(values = c("Preinduction" = "#feca8d", "1st week" = "#f1605d", "2nd week" = "#440f76")) +
  
  # Labels and theme
  labs(
    x = "Timepoint",
    y = "Social Preference %",
    fill = "Group",
    color = "Group"
  ) +
  theme_bw() +
  theme(
    text = element_text(size = 16),
    legend.position = "right"
  )

plot_raincloud
ggsave("3CT_sociability.svg", plot = plot_raincloud, width = 8, height = 6)
```



LLM

```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

novelty_cond$Week <- factor(novelty_cond$Week , levels = c("Preinduction","1st week post induction","2nd week post induction"))

levels(novelty_cond$Week)[levels(novelty_cond$Week) == "1st week post induction"] <- "1st week"
levels(novelty_cond$Week)[levels(novelty_cond$Week) == "2nd week post induction"] <- "2nd week"


#creating a linear model of the same variables as included in the anova
m_anova <-lmer(novelty_preference_relative ~  Week + (1|ID), data=novelty_cond)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week)

# for plot
emmeans_results_df <- as.data.frame(emmeans_results)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc
```




# assumption checks
```{r}
p_load(performance)

check <- check_model(m_anova)
check
```


#assumption checks


```{r}
#assumption check 1 - Normality of residuals (should be normally distributed) 
qqnorm(resid(m_anova))
qqline(resid(m_anova))
shapiro.test(residuals(m_anova)) 
#probs a lil' bit non-normal, thus, when calculating confidence intervals, used "boot" method which assumes non-normality
```

```{r}
#checks for distribution, compliments residual plot
hist(resid(m_anova)) 
```

```{r}
#Assumption 2 - Homoscedasticity (variance should be consistent) - passes
plot(fitted(m_anova), resid(m_anova))
#meets assumption, evenly distributed, thus, Homoscedasticity
```

```{r}
#When you plot residuals by ID, you're looking for:
#No systematic pattern across individuals.
#No obvious trends within IDs (e.g., all high or all low residuals).
#Ideally, residuals centered around zero with some variation.
boxplot(resid(m_anova) ~ novelty_cond$ID)
```

```{r}
#Assumption 3 - confidence intervals - passes
confint(m_anova, method = "boot")
#if CI includes zero between 2.5% and 97.5% then that effect is not statistically significant.

```


# Raincloud

```{r}
# Convert emmeans pairwise comparisons to data frame (BONFERRONI already applied in pwc)
pairwise_results_oft <- as.data.frame(pwc)

# Filter significant results (optional, not required for plotting)
sig_results_oft <- pairwise_results_oft %>%
  dplyr::filter(p.value < 0.05)

# Add significance labels
sig_labels_oft <- pairwise_results_oft %>%
  dplyr::mutate(
    label = dplyr::case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ "ns"
    )
  )

# View labeled results
sig_labels_oft
```


```{r}

library(dplyr)
library(tidyr)

# IMPORTANT: ensure this is a data.frame
pairwise_results_oft <- as.data.frame(pwc)

# pre-compute safe y-position reference (NEVER rely on df inside pipe)
y_max <- max(novelty_cond$novelty_preference_relative, na.rm = TRUE)

sig_labels_oft <- pairwise_results_oft %>%
  
  # split contrasts into groups
  separate(contrast, into = c("group1", "group2"), sep = " - ") %>%
  
  mutate(
    # significance labels (Bonferroni already in pwc)
    label = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01 ~ "**",
      p.value < 0.05 ~ "*",
      TRUE ~ "ns"
    ),
    
    # safe, automatic y positioning
    y.position = y_max * (1 + 0.10 * row_number())
  ) %>%
  
  # explicitly use dplyr::select to avoid masking issues
  dplyr::select(group1, group2, y.position, label)

sig_labels_oft

```


```{r}
p_load(ggpubr, ggdist)

plot_raincloud <- ggplot(novelty_cond, aes(x = Week, y = novelty_preference_relative, fill = Week, color = Week)) +
  
  # Half-violin, nudged left
  stat_halfeye(
    side = "left",
    adjust = 0.5,
    justification = 1.1,
    .width = 0,
    point_colour = NA,
    alpha = 0.6,
    scale = 0.5, #make plot narrower
    position = position_nudge(x = -0.05)
  ) +
  
#  # Boxplot
#  geom_boxplot(
#    width = 0.05,
#    outlier.shape = NA,
#    alpha = 0.7,
#    color = "black",
#    show.legend=FALSE,
#    position = position_nudge(x = -0.15)
#  ) +
  
  
  #connect IDs
  geom_line(
  aes(x = Week, y = novelty_preference_relative, group = ID),
  color = "gray50",
  linewidth = 0.5,
  alpha = 0.4
) +

 # Jittered points
  geom_jitter(
    width = 0.04,
    size = 2.0,
    alpha = 0.5,
  ) +
  
  #add means for SE bars
  geom_point(
    data = emmeans_results_df,
    aes(x = Week, y = emmean),
    color = "black",
    size = 3.0,
    inherit.aes = FALSE,
    position = position_nudge(x = -0.15)
  ) +
  
    
  #adding in model SE data
  geom_errorbar(
    data = emmeans_results_df,
    aes(x = Week, ymin = emmean - SE, ymax = emmean + SE),
    width = 0.10,
    inherit.aes = FALSE,
    linewidth = 0.6,
    position = position_nudge(x = -0.15)
  ) +
  
  
  # # Add in significant lines and asterisks
  # stat_pvalue_manual(
  #   sig_labels_oft,
  #   label = "p.adj.signif",
  #   tip.length = 0.02,
  #   size = 6,
  #   inherit.aes = FALSE
  # ) +
  
  stat_pvalue_manual(
    sig_labels_oft,
    label = "label",
    tip.length = 0.02,
    size = 6,
    inherit.aes = FALSE
  ) +

  # Manual colors
  scale_fill_manual(values = c("Preinduction" = "#feca8d", "1st week" = "#f1605d", "2nd week" = "#440f76")) +
  scale_color_manual(values = c("Preinduction" = "#feca8d", "1st week" = "#f1605d", "2nd week" = "#440f76")) +
  
  # Labels and theme
  labs(
    x = "Timepoint",
    y = "Social Novelty Preference %",
    fill = "Group",
    color = "Group"
  ) +
  theme_bw() +
  theme(
    text = element_text(size = 16),
    legend.position = "right"
  )

plot_raincloud
ggsave("3CT_social_novelty.svg", plot = plot_raincloud, width = 8, height = 6)
```


