---
title: "LDCT"
output:
  pdf_document: default
  html_document: default
  word_document: default
date: '2022-12-05'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load data
```{r}
# Loading
library(pacman)
p_load(readxl, tidyverse)
# xlsx files
df <- read_excel("EPM_longitudinal_data_4weeks.xlsx")


```



Removal of mouse 224 from dataset it turned it it was not double transgenic
```{r}
# #df$ID<- as.numeric(df$ID) #Change mouse number from class character to class numeric
df<-df[!(df$ID==429),]
df<-df[!(df$ID==374),]
# df<-df[!(df$ID==461),]
# df<-df[!(df$ID==463),]
# df<-df[!(df$ID==402),]

#remove TG animals to check single TG behaviour
# 
df<-df[!(df$ID==423),]
df<-df[!(df$ID==376),]
df<-df[!(df$ID==372),]
df<-df[!(df$ID==398),]
df<-df[!(df$ID==400),]
df<-df[!(df$ID==459),]



```


Time spent in dark chamber:
```{r}
#periphery_preference = time spent in periphery / time spent in center
 df<-  df %>% mutate(time_dark=df$`In zone Closed_arm1 cumulative`+df$`In zone Closed_arm2 cumulative`)
 df<-  df %>% mutate(time_light=df$`In zone Open_arm1 cumulative`+df$`In zone Open_arm2 cumaltive`+ df$`O1_round cumulative`+df$`O2_round cumulative`)
# 
# #relative time spent in light over dark
 df<-  df %>% mutate(light_preference=time_light/(time_light+time_dark)*100) 
#relative time spent in dark over light
df<-  df %>% mutate(dark_preference=time_dark/(time_light+time_dark)*100) 

```


plot


LMM

#light preference
```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(light_preference ~  Week + (1|ID), data=df)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc
```


#dark preference
```{r}
#Loading packages for linear mixed effects modelling
library(lmerTest)
library(lme4)

#creating a linear model of the same variables as included in the anova
m_anova <-lmer(dark_preference ~  Week + (1|ID), data=df)
summary(m_anova)

# Obtain estimated marginal means
emmeans_results <- emmeans(m_anova, ~ Week)

# Perform pairwise comparisons
pairwise_results <- contrast(emmeans_results, method = "pairwise")

# Display the results
summary(pairwise_results)

#bonferroni
pwc <- pairs(emmeans_results, adjust = "bon")
pwc
```


#dark preference
```{r}
cbbPalette <- c("#fcfdbf", "#feca8d", "#D67236", "#f1605d", "#9e2f7f", "#440f76")

p_load(ggbeeswarm, ggsignif)

df$Week <- factor(df$Week , levels = c("Preinduction","1st week post induction","2nd week post induction", "3rd week post induction"))

ggplot(df, aes(x = Week, y = dark_preference)) +
  geom_beeswarm(data = df, size = 2.5, aes(color = ID)) +
  ylab("Darkness preference") +
  xlab("Timepoint") +
  # geom_signif(comparisons = list(c("Preinduction", "3rd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 115,
  #             annotations = c("0.0052")) +
  #   geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 105,
  #             annotations = c("0.0128")) +
  #     geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
  #             map_signif_level = TRUE,
  #             y_position = 110,
  #             annotations = c("0.0046")) +
  geom_signif(comparisons = list(c("Preinduction", "3rd week post induction")),
              map_signif_level = TRUE,
              y_position = 115,
              annotations = c("0.1561")) +
    geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              y_position = 105,
              annotations = c("0.1204")) +
      geom_signif(comparisons = list(c("Preinduction", "2nd week post induction")),
              map_signif_level = TRUE,
              y_position = 110,
              annotations = c("0.1293")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = df, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))

```

```{r}
ggplot(df, aes(x = Week, y = light_preference)) +
  geom_boxplot(outlier.colour = "red", outlier.shape = 16, outlier.size = 2) +
  ylab("light preference %") +
  xlab("Timepoint") +
  theme_minimal(base_size = 16)
```

```{r}
levels(df$Week)[levels(df$Week) == "1st week post induction"] <- "1st week"
levels(df$Week)[levels(df$Week) == "2nd week post induction"] <- "2nd week"

ggplot(df, aes(x = Week, y = dark_preference)) +
  geom_beeswarm(data = df, size = 2.5, aes(color = ID)) +
  ylab("Darkness preference %") +
  xlab("Timepoint") +
  geom_signif(comparisons = list(c("Preinduction", "1st week")),
              map_signif_level = TRUE,
              y_position = 110,
              annotations = c("ns")) +
  geom_signif(comparisons = list(c("Preinduction", "2nd week")),
              map_signif_level = TRUE,
              y_position = 120,
              annotations = c("*")) +
  geom_signif(comparisons = list(c("1st week", "2nd week")),
              map_signif_level = TRUE,
              y_position = 130,
              annotations = c("**")) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = df, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 16))

```

Assumption checks for this approach

```{r}
#1. normality of residuals
# Fit the model
m_anova <- lmer(periphery_preference ~ Week + (1|ID), data = df, na.action = na.omit)

# Extract residuals
residuals <- residuals(m_anova)

# Q-Q plot for normality
qqnorm(residuals)
qqline(residuals)

#2. homoscedasticity (the same as homogeneity of variance)
# Residuals vs Fitted plot
plot(fitted(m_anova), residuals)
abline(h = 0, col = "red")


#3. Independence of Residuals

#Since this is a repeated measures design, you have accounted for within-subject correlation by including random effects. Ensure that there are no obvious patterns in the residuals versus time.

#4 Random Effects Structure:

#The random effects (e.g., intercepts and slopes for each subject) should be normally distributed. This can be checked by examining the distribution of the random effects.

# Extract random effects
ranef(m_anova)

# Check distribution of random effects
qqnorm(ranef(m_anova)$ID[,1])
qqline(ranef(m_anova)$ID[,1])

#5 Linearity:

#The relationship between the predictor variables and the response variable should be linear. This can be assessed by plotting the fitted values against the predictor variables.

# Plot fitted values vs predictors
plot(df$Week, fitted(m_anova))


```

# T-test

Data set where participants not completing all data collection time points are sorted out
```{r}
# Filter out so data is only for the first two weeks
TwoWeeks_df <- df %>% filter(Week == "Preinduction" | Week == "1st week post induction")

```

#Filter out subjects that did not complete
```{r}

# Determine the weeks each participant should have
required_weeks <- TwoWeeks_df %>% 
  distinct(Week) %>% 
  pull()

# Count the number of weeks each participant has data for
participant_week_counts <- TwoWeeks_df %>%
  group_by(ID) %>%
  summarize(week_count = n_distinct(Week))

# Identify participants with data for all required weeks
complete_participants <- participant_week_counts %>%
  filter(week_count == length(required_weeks)) %>%
  pull(ID)

# Filter the dataset to include only complete participants
df_complete <- TwoWeeks_df %>%
  filter(ID %in% complete_participants)

# View the new dataset
head(df_complete)
```

```{r}
df_complete$ID <- as.factor(df_complete$ID)
df_complete$Week <- factor(df_complete$Week , levels = c("Preinduction","1st week post induction"))

```


```{r}
#assumption checks
p_load(rstatix)
df_complete %>%
  group_by(Week) %>%
  identify_outliers(light_preference)

by(df_complete$light_preference, df_complete$Week, shapiro.test)
```

## light preference time
```{r}
# Perform paired t-test
t_test_result <- t.test(light_preference ~ Week, data = df_complete, paired = TRUE)

# Print the result of the t-test
print(t_test_result)

# Extract the p-value from the test result
p_value <- t_test_result$p.value
```
```{r}
# Function to convert p-value to significance annotations
get_significance_label <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}
```


```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
significance_label <- get_significance_label(p_value)

ggplot(df_complete, aes(x = Week, y = light_preference)) +
  geom_beeswarm(data = df_complete, size = 2.5, aes(color = ID)) +
  ylab("Light preference %") +
  xlab("Timepoint") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              annotations = significance_label) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = df_complete, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```


## light preference time
```{r}

#assumption checks
p_load(rstatix)
df_complete %>%
  group_by(Week) %>%
  identify_outliers(light_preference_freq)

by(df_complete$light_preference_freq, df_complete$Week, shapiro.test)
# Perform paired t-test
t_test_result <- t.test(light_preference_freq ~ Week, data = df_complete, paired = TRUE)

# Print the result of the t-test
print(t_test_result)

# Extract the p-value from the test result
p_value <- t_test_result$p.value
```
```{r}
# Function to convert p-value to significance annotations
get_significance_label <- function(p_value) {
  if (p_value < 0.001) {
    return("***")
  } else if (p_value < 0.01) {
    return("**")
  } else if (p_value < 0.05) {
    return("*")
  } else {
    return("ns")
  }
}
```


```{r}
p_load(ggbeeswarm, ggsignif)
# Get the significance label for the paired t-test p-value
significance_label <- get_significance_label(p_value)

ggplot(df_complete, aes(x = Week, y = light_preference_freq)) +
  geom_beeswarm(data = df_complete, size = 2.5, aes(color = ID)) +
  ylab("Distance (cm)") +
  xlab("Timepoint") +
  geom_signif(comparisons = list(c("Preinduction", "1st week post induction")),
              map_signif_level = TRUE,
              annotations = significance_label) +
  geom_crossbar(stat = "summary", fun = mean, width = 0.2, fatten = 1.5) +
  geom_errorbar(data = df_complete, stat = "summary", fun.data = mean_se, width = 0.1, position = "dodge") +
  theme(text = element_text(size = 13))
```
